Cloud Supercomputing

### 2019-04-08 writing/paper/paper.tex
Fix typos & other errors

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
\section{Introduction}
\subsection{The ExaHyPE-Engine}
The \exahype{} project (\url{www.exahype.eu}) develops an engine to solve arbitrary hyperbolic partial differential equations (\pde{}s) on upcoming exascale high performance computing architectures.
A user of the engine is provided with a simple code interface which mirrors the parts required to formulate a well posed Cauchy problem for a system of hyperbolic \pde{}s of first order.
The actual numerical scheme, domain decomposition and parallelisation techniques are left as black box to the user.
Internally the \exahypeengine{} implements a high order \aderdg{} scheme on Legendre basis nodes.
The mesh is incorporated from the Peano framework~\cite{weinzierl2011}, which is based on spacetrees for Cartesian meshes.
The framework provides a distributed (\mpi{}) and shared memory (Intel \textsc{tbb}) parallelization tailored to dynamic adaptive mesh refinement (\amr{})~\todo{LR: ref}.
As the ADER-DG scheme is highly volatile as soon as shocks develop we additionally provide \textit{a posteriori} finite volume limiting, recently developed by \citeauthor{dumbser2016simple} in~\cite{dumbser2016simple}.
We want to emphasize the recent work~\cite{charrier2018stop} by \citeauthor{charrier2018stop} which describes our communication-avoiding implementation of the whole numerical scheme.
\newcommand{\diffCoeff}{\varepsilon}%
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure)
  \end{pmatrix}
}%
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}
   \end{pmatrix}
}%
We \todo{LR: General introduction is missing. Where do we see our paper, what has been done before? Here we should also introduce and describe the ExaHyPE-Engine as base of the paper} solve the compressible Navier-Stokes equations in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantitites $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
The conserved quantities
\begin{equation} 
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$, the energy density $\QE$.
In \cref{eq:equation-set}, the rows are the conservation of mass, the conservation of momentum, the conservation of energy.

@@

We model the diffusivity of the flow by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right),
\end{equation}
with constant viscosity $\mu$.
The heat diffusion is governed by the coefficient
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R = \frac{\mu c_p}{\Pr},
\end{equation}
where the ratio of specific heats $\gamma$, the heat capacity at constant pressure $c_p$ and the Prandtl number $\Pr$ depend on the fluid.

Many realistic atmospheric flows can be described by a pertubation over a background state that is in hydrostatic equilibrium
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state, where the pressure gradient is exactly in balance with the gravitational source term.
The momentum equation is dominated by the background flow in this case.
This leads to numerical instabilities.
To avoid this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ into a sum of the background pressure $\backgroundPressure(z)$ and pertubation $\pressure'(\bm{x}, t)$.
We split the density $\Qrho = \backgroundRho + \Qrho'$ in the same manner and arrive at
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}

In the rest of the paper, we make the following contributions:
\begin{itemize}
\item
  We utilize an efficient and  explicit \aderdg{}-scheme with shared and distributed memory parallelization for our equation set.
\item We show an \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric.
\end{itemize}

@@

Rules for dynamic \amr{} can be stated by the user with element wise checks against a custom defined set of criteria. We use an element local error estimate based on the total variation of the numerical solution. 
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.\todo{If using muscl, add sentence about finite differences estimate}
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{delete}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{delete} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither refine nor delete all cells. \todo{LR: Was gibt es an compressible NS mit dynamic AMR, wie unterscheiden wir uns im Refinement Kriterium?}
This inequality holds for arbitrary distributions under the weak assumption that their mean $\mu$ and standard deviation $\sigma$ are finite~\cite{wasserman2004all}.

@@

\section{Results}
We use a diverse mix of benchmarking scenarios that consist of standard \textsc{cfd} scenarios, and atmospheric flows.

@@

Our atmospheric flow scenarios are described in terms of the potential energy
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o$~\cite{muller2010adaptive,giraldo2008study}.
For all remaining scenarions, we use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = \SI{10e5}{\Pa}, \qquad g = \SI{9.8}{m/s^2}.
\end{equation}

>>>>

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
\section{Introduction}
\subsection{The ExaHyPE-Engine}
The \exahype{} project (\url{www.exahype.eu}) develops an engine to solve arbitrary hyperbolic partial differential equations (\pde{}s) on upcoming exascale high performance computing architectures.
A user of the engine is provided with a simple code interface which mirrors the parts required to formulate a well posed Cauchy problem for a system of hyperbolic \pde{}s of first order.
The actual numerical scheme, domain decomposition and parallelisation techniques are left as a black box to the user.
Internally the \exahypeengine{} implements a high order \aderdg{} scheme on Legendre basis nodes.
The mesh is incorporated from the Peano framework~\cite{weinzierl2011}, which is based on spacetrees for Cartesian meshes.
The framework provides a distributed (\mpi{}) and shared memory (Intel \textsc{tbb}) parallelization tailored to dynamic adaptive mesh refinement (\amr{})~\todo{LR: ref}.
As the ADER-DG scheme is highly volatile as soon as shocks develop, we additionally provide \textit{a posteriori} finite volume limiting, recently\todo{Is 2016 recently? Originally developed in this paper, or just this special version?} developed by \citeauthor{dumbser2016simple} in~\cite{dumbser2016simple}.
We want to emphasize the recent work~\cite{charrier2018stop} by \citeauthor{charrier2018stop}, which describes our communication-avoiding implementation of the whole numerical scheme.
\newcommand{\diffCoeff}{\varepsilon}%
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure)
  \end{pmatrix}
}%
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}
   \end{pmatrix}
}%
We \todo{LR: General introduction is missing. Where do we see our paper, what has been done before? Here we should also introduce and describe the ExaHyPE-Engine as base of the paper} solve the compressible Navier-Stokes equations in the conservative form% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantitites $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
The conserved quantities
\begin{equation} 
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$ and the energy density $\QE$.
The rows of \cref{eq:equation-set} are the conservation of mass, the conservation of momentum and the conservation of energy.

@@

We model the diffusivity by the stress tensor
\begin{equation}
  \label{eq:stress-tensor}
  \stressT(Q, \nabla Q) =
  \mu
  \left(
  \left(\nicefrac{2}{3} \divergence{\Qv} \right) -
  \left( \gradient{\Qv} + \gradient{\Qv}^\intercal \right)
  \right),
\end{equation}
with constant viscosity $\mu$.
The heat diffusion is governed by the coefficient
\begin{equation}
  \label{eq:heat-conduction-coeff}
  \kappa = \frac{\mu \gamma}{\Pr} \frac{1}{\gamma - 1} R = \frac{\mu c_p}{\Pr},
\end{equation}
where the ratio of specific heats $\gamma$, the heat capacity at constant pressure $c_p$ and the Prandtl number $\Pr$ depend on the fluid.

Many realistic atmospheric flows can be described by a pertubation over a background state that is in hydrostatic equilibrium
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state, where the pressure gradient is exactly in balance with the gravitational source term.
The momentum equation is dominated by the background flow in this case.
This can lead to numerical instabilities.
To avoid this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ into a sum of the background pressure $\backgroundPressure(z)$ and pertubation $\pressure'(\bm{x}, t)$.
We split the density $\Qrho = \backgroundRho + \Qrho'$ in the same manner and arrive at
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}

In the rest of the paper, we make the following contributions:
\begin{itemize}
\item
  We utilize an efficient and  explicit \aderdg{}-scheme with shared and distributed memory parallelization for our equation set.
\item We show an \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric flows.
\end{itemize}

@@

The user can state dynamic \amr{} rules by supplying custom criteria that are evaluated pointwise.
We use an element local error estimate based on the total variation of the numerical solution. 
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.\todo{If using muscl, add sentence about finite differences estimate}
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{delete}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{delete} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither refine nor delete all cells. \todo{LR: Was gibt es an compressible NS mit dynamic AMR, wie unterscheiden wir uns im Refinement Kriterium?}
This inequality holds for arbitrary distributions under the weak assumption that their mean $\mu$ and standard deviation $\sigma$ are finite~\cite{wasserman2004all}.

@@

\section{Results}
We use a diverse mix of benchmarking scenarios that consist of standard \textsc{cfd} scenarios and atmospheric flows.

@@

Our atmospheric flow scenarios are described in terms of the potential energy
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o$~\cite{muller2010adaptive,giraldo2008study}.
We use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = \SI{10e5}{\Pa}, \qquad g = \SI{9.8}{m/s^2},
\end{equation}
for all remaining scenarios.
